Implementing Capon Adaptive Beamforming on the GPU for Real Time Cardiac Ultrasound Imaging

### 2012-09-11 HRUIonGPU/jrnl_article.tex
started on IUS article

%\IEEEPARstart{I}{ntroduction} to Capon adaptive beamforming, ultrasound imaging and GPU/CUDA.
\IEEEPARstart{C}{apon} adaptive beamforming \cite{Capon1969} applied to medical ultrasound imaging has recently shown promising results in several studies \cite{Synnevag2007, Austeng2008, Vignon2008, Viola, Mehdizadeh2012}. It has been shown that the method introduces some interesting trade-offs in addition to just improved resolution, like maintaining image resolution in combination with higher framerates, smaller probes, or deeper penetration \cite{Synnevag2009}. Despite these benefits, widespread use has been limited by the immense computational complexity associated with the method \cite{So2011}.  In this paper we introduce the first implementation of the Capon beamformer exhibiting real-time performance when applied in a typical cardiac ultrasound imaging setting. To accomplish this task we make use of the massive, and easily accessed, parallel processing power found in modern Graphics Processing Units (GPU's), combined with novel methods to reduce the computational complexity as the problem size grows.

@@

\subsection{Solving Multiple Small Linear Systems}
In the literature, when it comes to solving systems of linear equations on the GPU the focus has mostly been on large systems, and GPU libraries therefore often lack a comprehensive collection of batched solvers. The reason for this is two-fold, first the GPU needs to solve thousands of small systems in order to beat the low memory latency of the CPU, second there is a range of system dimensions that does not fit well with the GPU architecture, where only a small amount of near-core memory and registers are available. Because of this, it has been proven hard to optimize a manageable set of solvers that provides speedups compared with the CPU in every cases. It is therefore not given that solving on the GPU is faster than solving on the CPU in general. However, there is two arguments for doing solving on the GPU. First the CPU, which already runs a lot of algorithms, is offloaded. Second the transfer of several thousand covariance matrices from CPU-side to GPU-side is spared.

>>>>

%\IEEEPARstart{I}{ntroduction} to Capon adaptive beamforming, ultrasound imaging and GPU/CUDA.
\IEEEPARstart{C}{apon} adaptive beamforming \cite{Capon1969} applied to medical ultrasound imaging has recently shown promising results in several studies \cite{Synnevag2007, Austeng2008, Vignon2008, Viola, Mehdizadeh2012}. It has been shown that the method introduces some interesting trade-offs in addition to just improved resolution, like maintaining image resolution in combination with higher framerates, smaller probes, or deeper penetration \cite{Synnevag2009}. Despite these benefits, widespread use has been limited by the immense computational complexity associated with the method \cite{So2011}.  In this paper we introduce the first implementation of the Capon beamformer exhibiting real-time performance when applied in a typical cardiac ultrasound imaging setting. To accomplish this task we make use of the massive, and easily accessed, parallel processing power found in modern Graphics Processing Units (GPUs), combined with novel methods to reduce the computational complexity as the problem size grows.

@@

\subsection{Solving Multiple Small Linear Systems}
In the literature, when it comes to solving systems of linear equations on the GPU the focus has mostly been on large systems, and GPU libraries therefore often lack a comprehensive collection of batched solvers. The reason for this is two-fold, first the GPU needs to solve thousands of small systems in order to beat the low memory latency of the CPU, second there is a range of system dimensions that does not fit well with the GPU architecture, where only a small amount of near-core memory and registers are available. Because of this, it has been proven hard to optimize a manageable set of solvers that provides speedups compared with the CPU in every cases. It is therefore not given that solving on the GPU is faster than solving on the CPU in general. However, there is two arguments for doing solving on the GPU. First the CPU, which might already runs a lot of algorithms, is offloaded. Second the transfer of several thousand covariance matrices from CPU-side to GPU-side is spared.
